---
title: "gender_wage_gap_analysis"
format: html
---

# Q1
```{r setup, include = FALSE}
library(AER)
library(broom) 
library(car)
library(dplyr)
library(epiDisplay)
library(foreign)
library(ggplot2)
library(lmtest)
library(modelsummary)
library(readr)
library(readxl)
library(sandwich)
library(tidyverse)

finals <- read_csv("gender_wage_gap_microdata.csv")
```

# Q2
```{r}
finals <- finals |>
  mutate(private = if_else(worker_type %in% c("private, profit", "private, non profit"), 1, 0))
```

# Q3
```{r}
stats_table <- finals |>
  summarise(across(any_of(c("education","female","earnings","private","black")),
                   list(mean = ~mean(.x, na.rm = TRUE),
                        var  = ~var(.x,  na.rm = TRUE),
                        sd   = ~sd(.x,   na.rm = TRUE)),
                   .names = "{.col}_{.fn}")) |>
  pivot_longer(everything(),
               names_to = c("variable", ".value"),
               names_sep = "_")

```

# Q4
```{r}
finals <- finals |>
  mutate(female = as.integer(tolower(gender) == "female"))

m1 <- lm(earnings ~ female, data = finals)
modelsummary(m1, vcov = "HC1")
```

# Q5
Based on the regression model m1, Women earn about $22472 less per year than men, on average

# Q6
```{r}
male_mean   <- coef(m1)[1]
female_mean <- coef(m1)[1] + coef(m1)[2]
```

# Q7
```{r}
m2 <- lm(log(earnings) ~ female, data = finals)
modelsummary(list(m1,m2))
```

# Q8
```{r}
m2 <- lm(log(earnings) ~ female, data = finals)
m3 <- lm(log(earnings) ~ female + education, data = finals)

modelsummary(list(m2, m3))
```

# Q9
Yes—biased upward. 
m2	−0.282
m3	−0.341
Controlling for education reveals a larger (more negative) gender gap, the simpler model m2 understated the gap. 

# Q10
```{r}
beta_hat_1 <- -0.282
beta_hat_2 <- -0.341

difference <- beta_hat_1 - beta_hat_2
print("the difference of the betas between education and female is: ")
print(difference)
print("hence the difference is positive, and beta_hat_edcuation is + .102. the fraction for cov(x,y)/var(x) must then be positive. ")
```
Women in this sample have, on average, more schooling than men.

# Q11
```{r}
N        <- nrow(finals)
n_female <- sum(finals$female)
p        <- n_female / N
var_D    <- p * (1 - p)
cov_DX   <- 0.5784314 * var_D
cov_DX
```

# Q12
```{r}
p       <- sum(finals$female) / nrow(finals)
var_D   <- p * (1 - p)
mu_X    <- sum(finals$education) / nrow(finals)
var_X   <- sum( (finals$education - mu_X)^2 ) / nrow(finals)

cov_DX  <- 0.059 / 0.102 * var_D

rho_DX  <- cov_DX / sqrt(var_D * var_X)
rho_DX
```

# Q13
```{r}
cov(finals$education, finals$female)
cor(finals$education, finals$female)
```

# Q14
```{r}
# dummy variable.
finals <- finals |>
  mutate(black_ = as.integer(black == "black" | black == 1))

#models
m3 <- lm(log(earnings) ~ female + education,              data = finals)
m4 <- lm(log(earnings) ~ female * black_ + education,    data = finals)

modelsummary(list(m3, m4))
```

# Q15
non‑Black women earn about 34.1% less than non‑Black men, on average in , after controlling in logarithmic terms for education.

# Q16
```{r}
beta_hat_female_times_black_ <- 0.301
beta_hat_4_female <- -0.371

differnece1 <- beta_hat_4_female + beta_hat_female_times_black_
differnece1

```
Black women earn about 7% less than Black men, on average, after controlling for education.

# Q17
```{r}
error_beta_hat_female_times_black_ <- 0.060

t_stat <- beta_hat_female_times_black_ / error_beta_hat_female_times_black_
t_stat

print("t_stat greater than 1.96, significant")
```
The positive, significant test shows that the gender wage penalty is much smaller for Black workers than for non‑Black workers, and the difference is (about 0.30log‑points) and is statistically reliable. 