---
title: "in-labQuiz9"
format: html
---

#Q1
```{r}
#| message: false
#| warning: false
#| echo: true

library(tidyverse)
library(modelsummary)

df <- read_csv("covid_vac.csv") 
```

#Q2
```{r}
df <- df %>%
  mutate(covid.19.deaths = ifelse(is.na(covid.19.deaths), 0, covid.19.deaths))
```

#Q3
```{r}
m1 <- lm(covid.19.deaths ~ series_complete_pop_pct, data = df)

modelsummary(m1,
  statistic = "std.error",
  stars     = TRUE)
```

#Q4
```{r}
m2 <- lm(covid.19.deaths ~ series_complete_pop_pct + census2019, data = df)


modelsummary(list(m1, m2),
  statistic = "std.error",
  stars     = TRUE)

```

#Q5
##### The raw count of COVID deaths is mechanically higher in larger counties. Adding `census2019` isolates the *effect* of vaccination rates while holding the size of the at‑risk population constant, preventing spurious correlation that merely reflects county size instead of vaccination efficacy. Model1 – no population control	+21.98 deaths/ppt. Model2 – with population control	−1.52 deaths/ppt. County population accounts for 86% of the variance in the death counts that Model1 misses. Vaccination share by itself explains only ~6%. Median‑size county (≈26k people). Large county (90thpct, ≈209k people)

