---
ftitle: "education_earnings_walkthrough"
format: html
---

## 1 - adhoc Reggression
### 1.1. 1.2, 1.3 Creating the date frame
```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(modelsummary)
library(AER)
library(broom) 

wage       <- c(25, 15, 16, 16, 31)
education  <- c(12,  7, 13, 14, 15)


college<- ifelse(education > 12, 1, 0)   
# 1 if education > 12, else=0

df <- tibble(wage, education, college)
df
```

### 1.4, 1.5, 1.6 Mean wage by college graduation status

```{r regression}
## Wage_i = Bo + B1 * education_i + Îµ 

df %>%
  group_by(college) %>%
  summarise(mean_wage = mean(wage))

## B1_hat  = cov(Y,X) / var(x)
## Bo_hat  = Y_hat - b1_hat * mean(x)

beta1_hat <- cov(df$wage, df$education) / var(df$education)
beta0_hat <- mean(df$wage) - beta1_hat * mean(df$education)

beta1_hat; beta0_hat

```

### 1.7 Verify results using lm() AND present result using modelsummary table

```{r modelsummary}

m_ed <- lm(wage ~ education, data = df)
modelsummary(m_ed)

#`lm(wage~college)` generates the regression and `modelsummary(lm(wage~college))` generate the table

```
### 1.8 Interpret Results for the education coefficient. 

####  Each additional year of schooling is associated with an estimated $r round(beta1, 2) increase in hourly wage (on average), holding all else equal in this model.
#### Î²1_hat = $1.17 per hr per year of schooling). On average, every extra year of education is associated with a $1.17 increase in hourly wage. 

### 1.9 Consider Wage_i = B0 + B1 * College_i + Îµ
```{r wagesummary}
m_col <- lm(wage ~ college, data = df)
modelsummary(m_col)
```
#### Switching from non-college to college ( college = 0 â†’ 1 ) raises the predicted wage by coefficient on $ college. 

## 2 Real Data
### 2.1 Load the CPSSW Education dataset
```{r import_cps}
df <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/AER/CPSSWEducation.csv")
```
## 2.2 Regress earnings on education
```{r regress_cps}
ols <- lm(earnings ~ education, data = df)
modelsummary(ols)
```

## 2.3 Interpret the slope

#### B1_hat = education has a slope of 1.467
#### B0_hat = the predicted y-intercept  when education is equal to 0. 
#### x_i = years
#### earnings_hat = predicted earnings value
#### each additional year of schooling is associated with an average increase of about $1,467 in annual earnings. 

## 2.4 t-stat, p-value, AND 95 % confidence interval. 
```{r stats_cps}
beta1 <- 1.467
se1   <- 0.070
t_stat <- beta1 / se1
df_size     <- 2950 - 2
p_value <- 2 * pt(-abs(t_stat), df_size)
t_stat  
p_value

# Confidence Interval
ci_l   <- beta1 - qt(0.975, df_size) * se1
ci_u   <- beta1 + qt(0.975, df_size) * se1
data.frame(beta1, se1, t_stat, p_value, ci_l, ci_u)
ci_l
ci_u
```

## 2.5 Verify with summary()
```{r ols_cps}
summary(ols)
```

## 2.6 Using the values of ð›½0 and ð›½1 calculate for each observation the predicted earnings level.
```{r betas_cps}
# assuming your data frame is called 'data' and has a column 'Education'
beta0 <- -3.134
beta1 <-  1.467

df$earning_hat <- beta0 + beta1 * df$education

head(df[, c("education","earnings","earning_hat")])
```

## 2.7 Verify your calculation using the command predict.
```{r verification_cps}
df$earning_hat2 <- predict(ols)
all.equal(df$earning_hat, df$earning_hat2)
```

