---
title: "lab3"
format: html
---

## 2.1

```{r}
library(readr)
library(readxl)
library(dplyr)
library(ggplot2)
wage       <- c(34.6, 19.2, 13.7, 13.9, 19.2) 
education  <- c(16,   16,   12,   13,   16)  
```

## 2.2

```{r}
df <- data.frame(wage, education)
df
```

## 2.3

```{r}
experience <- c(2, 4, 1, 3, 5)
experience
```

## 2.4

```{r}
gender <- c("male", "female", "female", "female", "male")  # five values
gender
```

## 2.5

```{r}
wage_mean   <- mean(df$wage)
wage_sd     <- sd(df$wage)
wage_median <- median(df$wage)

edu_mean    <- mean(df$education)
edu_sd      <- sd(df$education)
edu_median  <- median(df$education)

wage_mean
wage_sd
wage_median

edu_mean
edu_sd
edu_median
```

## 2.6

```{r}
library(dplyr)  
summarise(df, average=mean(wage), standard_deviation=sd(wage), median=median(wage))

summarise(df, average=mean(education), standard_deviation=sd(education), median=median(education))
```

##2.7

```{r}
df %>% summarise(mean(wage))
df %>% summarise(mean(education))

df %>% summarise(sd(wage))
df %>% summarise(sd(education))

df %>% summarise(median(wage))
df %>% summarise(median(education))
```

## 3.1

```{r}
url <- "https://vincentarelbundock.github.io/Rdatasets/csv/AER/CPSSWEducation.csv"
cps_sw_edu <- read_csv(url, show_col_types = FALSE)

cps_sw_edu
```

## 3.2

```{r}
CPS <- read_excel('/Users/marcosfuerte/ECON42/ECON42_Week3/CPSSWEducation.xlsx')
CPS1 <- read_excel('/Users/marcosfuerte/ECON42/ECON42_Week3/CPSSWEducation.xls')
```

## 3.3

```{r}
df %>% summarise(
  mean_education = mean(education),
  var_education  = var(education),
  sd_education   = sd(education),
  mean_wage      = mean(wage),
  var_wage       = var(wage),
  sd_wage        = sd(wage)
)
```

## 4.1, 4.2, 5.1

```{r}

# brings in ggplot2, dplyr, readr, etc.
df1 <- read_csv('/Users/marcosfuerte/Downloads/CPSSWEducation.csv')

# 3. Inspect (optional) -----------------------------------
glimpse(df1)          # quick structure check
#> Rows: 300  (example)
#> Columns: rownames, age, gender, earnings, education

ggplot(df1, aes(education, earnings, colour = gender)) +
  geom_point(size = 2, alpha = .8) +
  geom_smooth(method = "lm", se = FALSE,  # remove se = FALSE if you want the 95% band
              linewidth = .7) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    title  = "Earnings vs. Education, coloured by Gender\nwith fitted linear trends",
    colour = NULL
  ) +
  theme_classic()

```

## 5.2

### Looking at your graph, who appear to be paid more ? Male workers, Female workers, or both earn similarly. Justify how you manage to figure this out.\[1pt\]

#### The slops of both lines for male and female appear equavalent. However Male workers appear to be paid more.

#### the men’s regression line is shifted upward—so, visually and statistically, men appear to earn more than women at every level of schooling in this sample.

## 5.3

### Looking at your graph, who appear to have the highest return to education? Male workers, Female workers, or both earn similarly. Justify how you manage to figure this out.\[1pt\]

#### Men tend to have the highest return to education.

```{r}
m_par   <- lm(earnings ~ education + gender,         data = df1)  # reduced
m_full  <- lm(earnings ~ education * gender,         data = df1)  # full

anova(m_par, m_full)   # F‑test for the extra interaction terms
```

#### That means neither group enjoys a visibly higher return—the payoff to an extra year of education appears essentially identical for men and women. (The formal test you ran backs this up: the interaction term that would let the slopes diverge was not significant, p=0.73.)

## 6.1

```{r}
df1$higher_ed<-(df1$education>12)
```

## 6.2

```{r}
df1$higher_ed[df1$education>12]<-1
df1$higher_ed[df1$education<=12]<-0
```

## 6.3

```{r}
df1 %>% group_by(higher_ed) %>% summarise(mean(education))
```
