---
title: "Lab 3 – Education and Earnings"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 2 – Real data

## 2.1 Load the education data

```{r load-data}
# Load directly from Vincent Arel-Bundock's Rdatasets repository
url <- "https://vincentarelbundock.github.io/Rdatasets/csv/AER/CPSSWEducation.csv"
edu <- read.csv(url)
str(edu)
summary(edu)
```

The variables of interest are **earnings** (annual labour income, 2004 $) and **education** (years of completed schooling).

## 2.2 Regress *earnings* on *education*

```{r regression}
ols <- lm(earnings ~ education, data = edu)
# install.packages("modelsummary") # if needed
library(modelsummary)
modelsummary(ols,
             statistic = "({statistic})",
             stars = TRUE,
             title = "Table 1. OLS regression of earnings on education")
```

## 2.3 Interpret the education coefficient

> The estimated slope coefficient, `r round(coef(ols)["education"],2)`, implies that **each additional year of schooling is associated with an average increase of about $`r format(round(coef(ols)["education"],0), big.mark = ",")` in annual earnings, ceteris paribus**.

(Replace wording/styling if you prefer.)

## 2.4 Manually compute *t*, *p*-value and 95% confidence interval

```{r manual-calculations}
beta1  <- coef(ols)["education"]
se1    <- summary(ols)$coefficients["education","Std. Error"]
t_stat <- beta1 / se1
df     <- ols$df.residual
p_val  <- 2 * pt(abs(t_stat), df, lower.tail = FALSE)
ci_l   <- beta1 - qt(0.975, df) * se1
ci_u   <- beta1 + qt(0.975, df) * se1
data.frame(beta1, se1, t_stat, p_val, ci_l, ci_u)
```

## 2.5 Verify with `summary()`

```{r verify-summary}
summary(ols)
confint(ols)
```

The analytically computed statistics should match the output above.

## 2.6 Predicted earnings for each observation

```{r predict-manual}
edu$earning_hat <- fitted(ols)   # same as predict(ols)
head(edu[, c("earnings","earning_hat")])
```

## 2.7 Verify using `predict()`

```{r predict-function}
edu$earning_hat2 <- predict(ols)
all.equal(edu$earning_hat, edu$earning_hat2)
```

Knit this R Markdown file to produce the required **HTML** report.

---

**Tip:** If you knit to HTML and the regression table does not appear, make sure the `modelsummary` package is installed and loaded.
