---
title: "beauty_premium_wage_analysis"
format: html
---

# Q1 
```{r, message = FALSE, warning = FALSE}
library(readr)
library(dplyr)
library(modelsummary)
library(ggplot2)
library(AER)
library(car)

beauty <- read_csv("beauty.csv", show_col_types = FALSE)

print(beauty)
```

#2
```{r}
beauty$looks_num <- match(beauty$looks, c("homely", "below average", "average", "above average", "handsome"))

datasummary(looks_num + educ + exper + wage ~ Mean + SD + Min + Max + N, data = beauty)
```

#3
```{r}
beauty$looks_num <- match(beauty$looks, c("homely","below average", "Average", "Above average","handsome"))

ggplot(beauty, aes(x = looks_num, y = log(wage))) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Part B
#4
```{r}
mod1 <- lm(log(wage) ~ looks_num, data = beauty)
mod2 <- lm(log(wage) ~ factor(looks), data = beauty)
modelsummary(
  list(mod1, mod2),  stars     = c("*" = .10, "**" = .05, "***" = .01))
```

#5
Model1 imposes a straight‑line relationship: every notch in appearance is worth the same 9% wage bump.
Model2 lets each category differ freely. It shows that the “penalty zone” (homely→below‑average) is steep, whereas the “reward zone” (above‑average→handsome) is flatter and statistically weak. In other words, looking unattractive hurts more than looking especially good helps.

Model‑fit is nearly identical (R² ≈1–2%), confirming that looks alone explains only a sliver of wage variation. Still, the factor model slightly improves the joint significance of looks (F = 4.4 vs.4.0).

#6
```{r}
lht(mod2,
    c("factor(looks)Average      = 0",
      "factor(looks)below average = 0",
      "factor(looks)handsome      = 0",
      "factor(looks)homely        = 0"))
```

#7
No—our regressions only uncover correlation, not causation.
To treat the 5‑point looks score as an exogenous “treatment” we must believe that nothing else, observed or unobserved, influences both appearance and pay. 

#8
```{r}
beauty$looks_num <- match(beauty$looks,
                          c("homely","below average","Average",
                            "Above average","handsome"))

m1 <- lm(log(wage) ~ looks_num,                         data = beauty)
m2 <- lm(log(wage) ~ looks_num + educ,                  data = beauty)
m3 <- lm(log(wage) ~ looks_num + educ + exper,          data = beauty)
m4 <- lm(log(wage) ~ looks_num + educ + exper + union + goodhlth, data = beauty)
m5 <- lm(log(wage) ~ looks_num + educ + exper + union + goodhlth + black + female + married, data = beauty)

modelsummary(
  list(m1, m2, m3, m4, m5),
  estimate  = "{estimate}{stars}",
  statistic = "({std.error})",
  stars     = c("*" = .10, "**" = .05, "***" = .01))

```

#9
Specification m5 is the preferred model because it offers the best statistical fit (highest Adj. R², lowest AIC/BIC) and retains all theoretically relevant controls, most of which are individually significant, while leaving the main “beauty premium” estimate both stable and highly significant.

#10
```{r}
m6 <- lm(lwage ~ looks_num + educ + exper + exper^2 + union + goodhlth + black + female + married, data = beauty)

modelsummary(
  m6,
  estimate  = "{estimate}{stars}",
  statistic = "({std.error})",
  gof_omit  = "IC|RMSE"
)
```
Because the quadratic is concave (β₁>0, β₂<0), the experience–earnings profile rises, flattens, then slowly declines. exper^ = -β₁/2*β₂ = (0.039 / (2*0.00059)) = 33 years
Adding the quadratic improves model fit (Adj. R² rises from 0.357 to ~0.374) while leaving the beauty premium (looks_num ≈ 0.059. 

#11
exper_star = 0.0389 / (2 * 0.000586) = 33.1911262799
33.1911262799; 33.2 years.


#12
```{r}
mod_int <- lm(
  log(wage) ~ looks_num * female        # interaction term
             + educ + exper + I(exper^2)
             + union + goodhlth
             + black + married,
  data = beauty
)

modelsummary(mod_int,
             estimate  = "{estimate}{stars}",
             statistic = "({std.error})",
             stars     = c("*" = .10, "**" = .05, "***" = .01))
```

#13
```{r}
mod_gender <- lm(log(wage) ~ looks_num * female + educ + exper + union + goodhlth + black + married, data = beauty)

modelsummary(
  mod_gender,
  estimate  = "{estimate}{stars}",
  statistic = "({std.error})",
  stars     = c("*" = .10, "**" = .05, "***" = .01))
```

#14
Interaction model: ln(wage) = β0 + β1 * looks_num + β2 * female + β3 * (looks_num × female) + Xγ

Effect for men (female = 0):  ln(wage) / d looks_num = β1 = 0.043 (s.e. 0.025)
Moving up one step on the 5‑point appearance scale raises male log‑wages by ≈4.4% [exp(0.043) − 1]. The coefficient is significant at the 10% level (*).

Effect for women (female = 1): β1 + β3 = 0.043 + 0.036 = 0.079
Each one‑step improvement in looks is associated with an ≈8.2% higher wage for women [exp(0.079) − 1].

Increment relative to men: β3 = 0.036 (s.e. 0.041), not statistically different from zero at conventional levels (p ≈ 0.38), so we cannot reject the null hypothesis that the beauty premium is the same for both sexes.


#15
No. The interaction term that captures the extra beauty premium for women (β₃=0.036,s.e.=0.041) yields a t‑statistic of about 0.88 and a p‑value around 0.38—far above conventional 10%,5% or 1% significance thresholds. Hence we cannot reject the null hypothesis that the beauty effect is the same for men and women; any apparent difference in the point estimates is not statistically significant.


#16
```{r}
beauty <- beauty %>%
  mutate(
    looks_lower = tolower(looks),
    handsome    = ifelse(looks_lower == "handsome", 1, 0),
    homely      = ifelse(looks_lower == "homely",   1, 0)
  ) %>%
  select(-looks_lower)

table(beauty$handsome, beauty$looks)
table(beauty$homely,   beauty$looks)
```

#17
```{r}
m_extremes <- lm(
  log(wage) ~ handsome + homely +
              educ + exper + I(exper^2) +
              union + goodhlth +
              black + female + married,
  data = beauty
)

modelsummary(
  m_extremes,
  estimate  = "{estimate}{stars}",
  statistic = "({std.error})",
  stars     = c("*" = .10, "**" = .05, "***" = .01)
)
```
The overall “beauty effect” seen in earlier regressions is driven almost entirely by the wage penalty imposed on workers rated homely; the reward for being rated handsome is modest and statistically fragile. In short, unattractiveness hurts earnings more than exceptional attractiveness helps.



