---
title: "Econ 42Lab2.R"
format: html
---

#### This line will setup the current working directory
```{r SETTINGUP, include=FALSE}
setwd("/Users/marcosfuerte/ECON42/ECON42_Week2")
```

Using the libraries in order to import the modules
```{r Libraries, message=FALSE, warning=FALSE, error=FALSE}
## SETTING UP ## 

library(AER)
library(foreign)
library(readr)
```

#### Running Various calculations for the variables of X and Y; This chunk uses echo =FALSE 

```{r calc}
## CALCULATOR ##
#| echo: false
X <- c(2,6,7,9)
mean_X <- mean(X)
var_X <- var(X)
sd_X <- sd(X)

print(mean_X)
print(var_X)
print(sd_X)

summary(X)

Y <- c(2,6,7,9,15,18,20, 8792847, 72398145)
median_Y <- median(Y)
quantile(Y, 0.1)    # 10th percentile
quantile(Y, 0.25)   # 25th percentile

# Y is redefined
Y <- c(2,2,4,0)
cov_XY <- cov(X, Y)
cor_XY <- cor(X, Y)

summary(Y)

table <- table(X, Y)
```
#### This chunk is the dataframing process to display the variables of X and Y as tables. We also called in the files from the web using library(foreign) and to view files for .csv with library(readr); This chunk does not use echo =FALSE; View() cannot be rendered so I commented out the function.
```{r dataframing}
## GENERATE ##
X <- c(1,2,3,4)
Y <- c(9,8,7,6)
  
df<-data.frame(X,Y)
#View(df)

# url <- "http://fmwww.bc.edu/ec-p/data/stockwatson/caschool.dta"
# mode = "wb"; write binary
# download.file(url, destfile = "caschool.dta", mode = "wb")

# read.dta
df2<-read.dta("http://fmwww.bc.edu/ec-p/data/stockwatson/caschool.dta")
# View(df2)

# .csv file using library(readr)
df3 <- read_csv("caschool.csv")
# View(df3)
```
